---
title: "Lab 1"
author: "Dillon Niederhut"
date: "August 26, 2014"
output: ioslides_presentation
---

## Review introduction

For this course, you are expected to already have a mastery of algebra. We do not anticipate this to be a problem, but just to be safe, we will begin by reviewing the math concepts that you are expected to know.

## Order of operations

1. Evaluate expressions inside parentheses
2. Evaluate powers left to right
3. Evaluate multiplication/division left to right
4. Evaluate addition/subtraction left to right

## Try these by hand

- 4 + 2 * 3
- 32 * 4^2^
- (8 - 4)^2^/4

## Now type them into the console in R

```{r}
4 + 2 * 3
32 * 4 ** 2
((8 - 4) **2)/4
```

## Lines in Cartesian coordinates

A line is composed of two parts: a slope and an intercept

1. A slope is a coefficient attached to a variable, like 2X. You can calculate a slope by dividing the change in Y by the change in X.

2. An intercept is a constant value like 4. You can calculate an intercept by finding where a line crosses the Y axis.

## Imagine you have two data points (1,6) and (2,8)

The slope is calculated as `(Y2-Y1)/(X2-X1)`:
```{r}

m <- (8-6)/(2-1)
print(m)
```

and the intercept is calculated as `Y1 - m*X1`:
```{r}

b <- (6 - 1 * m)
print(b)
```

## Which gives you a line that looks like this:

```{r, fig.height = 4, fig.width = 7, echo = FALSE}

library(ggplot2)
x <- c(1,2)
y <- c(6,8)
data <- data.frame(x,y)
qplot(data = data, x, y) + geom_abline(intercept = 4, slope = 2)

```

## Now you try

- Calculate the slope and intercept of the line crossing through these two points: (4,6) (6,4)
- Check your answer by calculating this in R

```

data <- data.frame(x = c(4,6), y = c(6,4))
summary(lm(y ~ x, data))
```

## Summation

Summation notation is a shorthand for when you are adding a bunch of similar things together. They are evaluated like this:

```

 10
Sigma i = 1 + 2 + 3 + 4 + 5
 i = 1
```

## Try calculating these sums

```

 5
Sigma i^2^
 1


 5
(Sigma i)^2^
 1
```