---
title: "Lab 2"
author: "Dillon Niederhut"
date: "September 9, 2014"
output: ioslides_presentation
---

## Don't Panic!

```{r, echo=FALSE}
library(ggplot2)
load('survey_data_cleaned.Rda')
```


- We're going to start with simple things

- These simple things build on each other

- Make sure you are keeping up with the readings!

## Why plot stuff?

```{r fig.width=9, fig.height=5, echo=FALSE}
library(png)
library(grid)
img <- readPNG('/Users/dillonniederhut/Dropbox/Psych101/explot.png')
 grid.raster(img)
```

## Useful beginner steps in R (1)

- You can get a sense od the structure of your data using:
```{r}
str(data)
```

## Useful begiiner steps in R (2)

- You can get a sense of the properties of your data using:
```{r}
summary(data)
```

## Plotting is easy in R

```{r, fig.height = 4, fig.width = 7}
qplot(data = data, x = Q10_mother)
```

## Plotting is flexible in R

```{r, fig.height = 4, fig.width = 7}
ggplot(data = data, aes(x = Q10_mother, y = Q9_father)) + geom_jitter() + 
  stat_smooth(method = 'lm')
```

## Getting set up

- Login to your computers with:

- ID: !tmfpsych101

- Password: c@1thinking

- Start Rcommander with ```library(Rcmdr)```

- Instead of ```ClassFall14$inlove_factor```... you are going to use:
```data$Q8_fallinlove_factor <- factor(data$Q8_fallinlove)```
